04-analisis

### 2024-10-02 03-methods.qmd
template mdpi
agregue un nuevo qmd para el template mdpi
Update submit mdpi FINAL
@juancarloscastillo
coments methods & analisis

---
author: "Equipo Edumer"
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/apa6.csl"
---

@@

The First Study of Civic Education in Chile, conducted by the Agency for Quality Education of the Ministry of Education in 2017, serves as the main data source. This database is composed by a civic knowledge test score and a series of items batteries measuring different aspects of citizenship. The target population includes 8th-grade students in 242 schools nationwide, as well as their parents and teachers. Initially, the database contains 8,589 student observations and 6,770 parent observations. To ensure higher data quality and considering the survey's unit of analysis, we removed 171 student cases and 79 parent cases that exhibited repetitive and careless response patterns [@gottfried_autocorrelation_2022]. Additionally, we utilized data from the System for Measuring the Quality of Education (SIMCE) conducted by the Ministry of Education in 2017, which provides school-level information such as administrative dependence, socioeconomic classification, and results obtained in standardized mathematics and language tests. After processing the variables and removing missing cases, the final database for this study includes a two-level stratified sample composed of 5,047 students and parents (level 1) nested within 231 schools (level 2).

@@

**Market Justice Preferences**: The dependent variables in this study are market justice preferences. This construct is measured through three variables that address the degree of justification regarding whether access to social services in pensions, education, and health should be income conditional. Justification of inequality in health is measured by the item: "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes?" The same question is asked for education and pensions. In all cases, respondents indicate their preferences on a Likert scale ranging from "completely disagree" (1) to "completely agree" (4). Additionally, we include a summarized indicator of "market justice preferences", measured by an average index across these items (α = 0.86), with values ranging from 1 to 4, where higher values represent greater preferences for market justice (see @tbl-desc-dependientes). We analyzed these items independently as well as by the average index.

@@

This study focuses on two school-level characteristics: socioeconomic status and academic performance. Socioeconomic status is assessed using the Ministry of Education's classification, measured as an ordinal item with five categories ranging from "low" (1) to "high" (5). Academic performance is measured using the school's results in the SIMCE (System of Measurement of Educational Quality) standardized tests, which are administered yearly at different educational levels in areas such as language and mathematics. These results are categorized as "low," "medium," and "high." The contextual level items used, response categories, and their frequencies are detailed in @tbl-desc-school.

@@

The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open Science (OSF), the access to the document is available at this [link](https://doi.org/10.17605/OSF.IO/UFSDV). The statistical analysis for this research was conducted using the *lme4* package in R version 4.1.3.

>>>>

---
author: "Equipo EDUMER"
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
---

@@

The First Study of Civic Education in Chile, conducted by the Agency for Quality Education of the Ministry of Education in 2017, is the primary data source. This database comprises a civic knowledge test score and a series of items that measure different aspects of citizenship. The target population includes 8th-grade students in 242 schools nationwide. Initially, the database contains 8,589 student observations.To ensure higher data quality and considering the survey's unit of analysis, we removed 171 student cases that exhibited repetitive and careless response patterns [@gottfried_autocorrelation_2022]. Additionally, we utilized data from the System for Measuring the Quality of Education (SIMCE) conducted by the Ministry of Education in 2017, which provides school-level information such as administrative dependence, socioeconomic classification, and results obtained in standardized mathematics and language tests. After processing the variables and removing missing cases, the final database for this study includes a two-level stratified sample composed of 5,047 students (level 1) nested within 231 schools (level 2).

@@

**Market Justice Preferences**: The dependent variables in this study are market justice preferences. This construct is measured through three variables that address the degree of justification regarding whether access to social services in pensions, education, and health should be conditional on income. The Justification of inequality in health is measured by the following item: "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes?" The same question is asked for education and pensions. In all cases, respondents indicate their preferences on a Likert scale ranging from "completely disagree" (1) to "completely agree" (4). Additionally, we include a summarized indicator of "market justice preferences", measured by an average index across these items (α = 0.86), with values ranging from 1 to 4, where higher values represent stronger preferences for market justice (see @tbl-desc-dependientes). We analyzed these items independently and by the average index.

@@

This study focuses on two school-level characteristics: socioeconomic status and academic performance. Socioeconomic status is assessed using the Ministry of Education's classification, measured as an ordinal item with five categories ranging from "low" (1) to "high" (5). Academic performance is measured using the school's results in the SIMCE (System of Measurement of Educational Quality) standardized tests, administered yearly at different educational levels in language and mathematics. These results are categorized as "low," "medium," and "high." The contextual level items used, response categories, and their frequencies are detailed in @tbl-desc-school.

@@

The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open Science (OSF), and access to the document is available at this [link](https://doi.org/10.17605/OSF.IO/UFSDV). The statistical analysis for this research was conducted using the *lme4* package in R version 4.1.3.

### 2024-10-02 04-analisis.qmd
template mdpi
agregue un nuevo qmd para el template mdpi
coments methods & analisis

---
author: "Equipo Edumer"
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
---

@@

Model 1 in @tbl-ordinal-reg for the three dependent variables of justice in differential access to health, pensions, and education demonstrates a similar trend regarding the effect of the perception of meritocracy on these items, with all effects being statistically significant (*p* \< 0.05). Regarding the variables of social meritocracy, the three variables---talent, effort, and deservingness---show that as these increase, the justification for differentiated access to pensions, education, and health also increases. In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases; conversely, as perceived school effort increases, the justification decreases, holding all other variables constant.

Additionally, Model 1 examines family status through parents' educational level and the number of books at home. For health, pensions, and education, a negative but statistically non-significant association with parents' education is observed. However, having more than 25 books at home shows a negative and statistically significant effect on market justice in health and pensions, suggesting less justification for differentiated access compared to those with fewer books.

Model 2 in @tbl-ordinal-reg incorporates school-level variables. The results show that high socioeconomic level schools show greater preferences for market justice in health, pensions, and education compared to low socioeconomic level schools, with statistically significant results (*p* \< 0.05). Moreover, and contrary to our hypothesis, schools with higher academic performance show negative and statistically significant effects across the three items.

@@

@tbl-lineal-reg presents the results of linear mixed-effects models for the market justice preference index. Model 1 shows that higher perceived societal meritocracy, through effort, talent, and deservingness, is positively and significantly associated with market justice preferences. School meritocracy perceptions yield mixed effects: perceived talent increases market justice preference, while perceived effort decreases it, both statistically significant (*p* \< 0.001). Model 2 includes family status, finding no statistically significant effects from parents' education or the number of books at home. Regarding school characteristics, Models 3 and 4 introduce socioeconomic level and academic performance. The results of Model 3 suggest a statistically significant relationship between school socioeconomic levels and market justice preferences, specifically those schools with high socioeconomic levels ($\beta$ = 0.37) compared to low socioeconomic levels. However, Model 4 indicates that schools with medium ($\beta$ = -0.11) and high ($\beta$ = -0.26) academic performance present lower market justice preferences compared to low-performing schools, with these effects being statistically significant when controlling for other variables.

@@

@tbl-interact displays the interaction terms related to our hypotheses. The results suggest that students' family background moderates the relationship between their meritocratic perceptions and their justification of market allocation in social services. Specifically, the relationship between social effort and market justice preferences is less positive for students whose parents have university or postgraduate education. Lower-status students (measured by parental education) prefer more market justice when they strongly adhere to meritocratic perceptions of deservingness. A similar moderating effect is seen with family cultural capital: the relationship between effort and market justice preferences is less positive for students with more than 25 books at home.

>>>>

---
author: "Equipo EDUMER"
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
---

@@

Model 1 in @tbl-ordinal-reg for the three dependent variables of justice in differential access to health, pensions, and education demonstrates a similar trend regarding the effect of the perception of meritocracy on these items, with all effects being statistically significant (*p* < 0.05). Regarding the variables of social meritocracy, the three variables—talent ($\beta$ = 0.16), effort ($\beta$ = 0.11), and deservingness ($\beta$ = 0.51)—show that as these increase, the justification for differentiated access to pensions, education, and health also increases. In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases ($\beta$ = 0.23); conversely, as perceived school effort increases, the justification decreases ($\beta$ = -0.26), holding all other variables constant.

Additionally, Model 1 examines family status through parents’ educational level and the number of books at home. For health, pensions, and education, a negative but statistically non-significant association with parents’ education is observed. However, having more than 25 books at home shows a negative and statistically significant effect on market justice in health and pensions ($\beta$ = -0.13; *p* < 0.001), suggesting less justification for differentiated access compared to those with fewer books.

Model 2 in @tbl-ordinal-reg incorporates school-level variables. The results show that high socioeconomic level schools show greater preferences for market justice in health, pensions, and education compared to low socioeconomic level schools, with statistically significant results ($\beta$ = 0.71; *p* < 0.05). Moreover, and contrary to our hypothesis, schools with higher academic performance show negative and statistically significant effects across the three items ($\beta$ = -0.56; *p* < 0.001).

@@

@tbl-lineal-reg ppresents the results of linear mixed-effects models for the market justice preference index. Model 1 shows that higher perceived societal meritocracy, through effort ($\beta$ = 0.08), talent ($\beta$ = 0.07), and deservingness ($\beta$ = 0.2), is positively and significantly associated with market justice preferences (*p* < 0.001). School meritocracy perceptions yield mixed effects: perceived talent increases market justice preference ($\beta$ = 0.1), while perceived effort decreases it ($\beta$ = -0.12), both statistically significant (*p* < 0.001). Model 2 includes family status, finding no statistically significant effects from parents’ education or the number of books at home. Regarding school characteristics, Models 3 and 4 introduce socioeconomic level and academic performance. The results of Model 3 suggest a statistically significant relationship between school socioeconomic levels and market justice preferences, specifically those schools with high socioeconomic levels ($\beta$ = 0.37; *p* < 0.01) compared to low socioeconomic levels. However, Model 4 indicates that schools with medium ($\beta$ = -0.11; *p* < 0.01) and high ($\beta$ = -0.26; *p* < 0.001) academic performance present lower market justice preferences compared to low-performing schools, with these effects being statistically significant when controlling for other variables.

@@

@tbl-interact displays the interaction terms related to our hypotheses. The results suggest that students’ family background moderates the relationship between their meritocratic perceptions and their justification of market allocation in social services. Specifically, the relationship between social effort and market justice preferences is less positive for students whose parents have university or postgraduate education ($\beta$ = -0.079; *p* < 0.05). Lower-status students (measured by parental education) prefer more market justice when they strongly adhere to meritocratic perceptions of deservingness ($\beta$ = -0.118; *p* < 0.01). A similar moderating effect is seen with family cultural capital: the relationship between effort and market justice preferences is less positive for students with more than 25 books at home ($\beta$ = -0.081; *p* < 0.01).
